---
title: "MakeoverMondays-Christmas Spend"
author: '@deuxlappins'
date: "12/22/2019"
output:
  html_document: default
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(knitr)
library(ggrepel)
```

## Introduction
This document will present visualization based on the Christmas Spending data in www.makeovermondays.co.uk for week 52.
The data includes actual spending for the year 2018 and the projected spending for 2019 for Christmas for UK and Rest of Europe.

## Loading the data 
To begin with let us load the data from the excel file and display the data and summary statistics.

```{r echo=TRUE}
christmasspending<-readxl::read_xlsx("../../Christmas_spending/Data/Christmas spending.xlsx")

knitr::kable(christmasspending)

```

## Transforming and cleaning the data
The column names are slightly unwieldy and the total is not really a category. Editing column names.

```{r echo=TRUE}
christmasspending<- christmasspending %>% rename(SpendingAmount=`Amount in £`)
christmasspending$Year <- as.factor(christmasspending$Year)
cmasNonTotal <- christmasspending %>% filter(Category != "Total")
knitr::kable(head(christmasspending))

```

## simplyfying this for easier viewing
To make for better viewing.
```{r echo=TRUE}

kable(cmasNonTotal %>% pivot_wider(names_from = "Year",values_from = "SpendingAmount"))
```

## Visualizing using ggplot

```{r echo=TRUE}
myplot<-ggplot(data = cmasNonTotal, 
               mapping = aes(x=Year,y=SpendingAmount,group=Category)) +
  geom_line(aes(colour=Category,alpha=.5),size=2) +
  geom_point(aes(colour=Category,alpha=.5),size=4) +
  geom_text_repel(data = cmasNonTotal%>%filter(Year=="2018")
                  ,color="white"
                  ,nudge_x = -.45
                  ,mapping = aes(label=paste0(Category," ", "£",as.character(SpendingAmount)),
                                 hjust = "left", 
                                 size = 1)) +
  geom_text_repel(data = cmasNonTotal%>%filter(Year=="2019")
                  ,color='white'
                  ,nudge_x = .25
                  ,mapping = aes(label=paste0( "£",as.character(SpendingAmount)),
                                 hjust = "left", 
                                 size = 1 
                                 
                  ))+
  
  
  scale_x_discrete(position = "top",limit=c("2018","2019")) +
  
  
  theme_bw() +
  
  theme(legend.position = "none") +
  theme(panel.border = element_blank()) +
  
  theme(axis.title.y     = element_blank()) +
  theme(axis.text.y      = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.grid.minor.y = element_blank()) +
  # Remove a few things from the x axis and increase font size
  theme(axis.title.x     = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(axis.text.x.top      = element_text(size=12)) +
  # Remove x & y tick marks
  theme(axis.ticks       = element_blank()) +
  facet_wrap(.~Region) +
  theme(strip.background = element_rect(
    color="white", fill="white"
  )) +
  
  theme(strip.text = element_text(face = "bold")) +
  
  #   #titles
  labs( title = "Comparision of Christmas spending between 2018 and 2019 across UK and Europe")

grb<-ggplotGrob(myplot)

grb$grobs[[2]]$children[[1]]$children[[1]]$gp$fill<-"#003399"
grb$grobs[[3]]$children[[1]]$children[[1]]$gp$fill<-"#00247D"
# # texts to match color
 grb$grobs[[12]]$grobs[[1]]$children[[2]]$children[[1]]$gp$col<-"#003399"
 grb$grobs[[13]]$grobs[[1]]$children[[2]]$children[[1]]$gp$col<-"#00247D"
plot(grb)
```

## Saving the Plot
```{r echo=TRUE}
ggsave("../../Christmas_spending/Outputs/Slopegraph.png",plot = grb)
```
